Sexual dimorphism in trait variability and its eco-evolutionary

### 2020-10-28 scripts/MouseSexDiffVar_Oct_2020.Rmd
reverting to rbind, with R version 4.0 / associated dplyr version; Changed Figure names to match the manuscriot files. Added doi for the knitted version

---
title: 'Sex and Power: Sexual dimorphism in trait variability and its eco-evolutionary
  and statistical implications'
author: Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky, Robert  Brooks,
  Will Cornwell, Daniel Falster, Malgorzata Lagisz, Jeremy Mason, Alistair Senior,
  Daniel Noble, & Shinichi Nakagawa
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 6
    toc_float: yes
  bookdown::pdf_document2:
    toc: yes
    toc_depth: 6
    number_sections: true
    theme: flatly
  word_document:
    toc: yes
    toc_depth: '4'
  html_notebook:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
  bookdown::word_document2:
    toc: yes
    toc_depth: 6
    number_sections: true
    theme: flatly
subtitle: Electronic Supplementary Material
always_allow_html: yes
---

@@

```{r FigG1, echo = FALSE, fig.width = 4, fig.height = 8, fig.cap="Workflow of data process and meta-analysis"}

@@

```{r FigCor, fig.height=3, message = FALSE, fig.cap= "Mean-variance relationships (log(Mean) vs log(SD, standard deviation)) across all traits for males (A) and females (B)."}

@@

# Meta-analysis results
## Preparing: Figure 3 
### Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. Preparation of subplot 3A

@@

## Main Manuscript: Figure 3

```{r Fig3, fig.cap = "Panel A shows the numbers of traits across functional groups that are either male-biased (blue-green) or female-biased (orange-red), as calculated in Figure 1.1D. Panel B shows effect sizes and 95% CI from separate meta-analysis for each functional group (step H in Figure1.1). Both panels represent results evaluated across all traits. Traits that are male biased are shown in blue, whereas female bias data is represented in orange."}
Fig4 <- ggarrange(malebias_Fig2_alltraits, Metameta_Fig3_alltraits,  nrow = 2, align = "v", heights = c(10, 9), labels = c("A", "B"))
Fig4
```

@@

plot3.male.meta.CVR.b <- bind_rows(plot3.male.meta.CVR.b, add.row.hearing)

@@

overall4.male.sig <- bind_rows(lnCVR.ci, lnRR.ci) %>% 

@@

plot3.female.meta.CVR <- bind_rows(plot3.female.meta.CVR, plot3.female.meta.CVR.all)

@@

plot3.female.meta.CVR.b <- bind_rows(plot3.female.meta.CVR.b, add.row.hearing)

@@

overall4.female.sig <- bind_rows(lnCVR.ci, lnRR.ci) %>% 

@@

overall4.male.sigS <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

overall4.female.sigS <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) # Note: bind_rows may need to be replaced by rbind

@@

overall4.male.sigS <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)
# may require: 	 %>% mutate(value = as.numeric(value), ci.low = as.numeric(ci.low), ci.high = as.numeric(ci.high))

@@

## Figure 5.1
Figure \@ref(fig:FigS1) is similar to Figure 3 in the main article except that lnVR has been added to the plot for comparison with lnCVR and lnRR.

```{r FigS1, fig.height = 8, fig.width = 7, fig.cap=" Numbers of either male (blue-green bars) or female (orange-red bars) biased traits  (Panel A) across functional groups, this time for lnCVR (left hand side), lnVR (middle) and lnRR (right hand side). Panel B shows effect sizes from separate meta-analysis for each functional group, and Panel C contains results of heterogeneity analyses. All three panels represent results evaluated across all traits."}

@@

FigS1 <- ggarrange(malebias_FigS1_alltraits, Metameta_FigS1_alltraits, HeteroS1, nrow = 3, align = "v", heights = c(1.1, 1, 1), labels = c("A", "B", "C"))
FigS1
#ggsave("FigS1_OverallResults.pdf", plot = FigS1, width = 6, height = 7)
```

@@

#### Methods
Among the 147 traits (i.e. with ‘first-order’ meta-analytic results), we found 31 traits to be significantly male-biased and 30 female-biased for lnCVR (47 male-biased and 39 female- biased for lnRR; 74 male and 55 female-biased for lnVR). We used this statistical significance to categorize traits into two groups: male-based and female-biased traits. For each set of these sex-biased traits, we conducted second-order meta-analyses with trait as the unit of analysis in the same way as those using lnRR, lnVR and lnCVR (Figure \@ref(fig:FigG1)J).  As sensitivity analyses, we also employed another way of separating male- and female-biased traits. We quantified traits that were more than 10% biased towards either of the sexes, regardless of significance (N[male-biased] = 31, & N[female-biased] = 36 for lnCVR,  N[male-biased] = 56, & N[female-biased] = 57 for lnVR , N[male-biased] = 42, & N[female-biased] = 33  for lnRR; see Supplemental Figure S\@ref(fig:FigS2), Supplemental Table 7.3).

#### Results
Having established that different sex-biases were present across trait functional groups, we explored specific patterns in male and female bias by categorizing traits into either male or female-biased, and then, quantifying the degree of sex bias. To do this, we focused on only those traits that were significantly biased towards one sex (as indicated by “CI not overlapping 0” in Figure S\@ref(fig:FigS2). Selecting significantly sex-biased traits reduced the number of traits substantially (Figure S\@ref(fig:FigS2)A; compared to all traits, Figure \@ref(fig:Fig3) and Figure \@ref(fig:FigS1)), and returned similar numbers of traits with either significant male- or female-bias for lnCVR (Figure S\@ref(fig:FigS2)A).

Looking only at traits with significant male-bias, trait variability tended to be higher in males than in females overall (11.77% male bias, lnCVR = 0.111, CI = [0.093 to 0.129]; compared to 8.3% female-bias, lnCVR = -0.087, CI = [-0.109 to (-0.064)]; Figure S\@ref(fig:FigS2)B). The degree of sex-bias varied across the functional trait groups, ranging from 4.6 % (Physiology, lnCVR = -0.047, CI = [-0.064 to (-0.030)] to 15.13 % (Eye morphology, lnCVR = -0.164, CI = [-0.249 to (-0.250)] in females, and from 8.6 % (Hematology, lnCVR = 0.082, CI = [0.048 to 0.117]) and 16.66 % (Heart, lnCVR = 0.154, CI = [0.085 to 0.223]) in males (Figure S\@ref(fig:FigS2)B).

To put sex biases (sexual dimorphisms) in variability into perspective, we also quantify the degree of sexual dimorphism in significantly sex-biased traits (i.e., the extent of sexual dimorphisms). There were more traits that were significantly biased in mean trait value in males compared to females (Figure S\@ref(fig:FigS2)A, lnRR). These traits were 11.73 % biased towards males (Figure S\@ref(fig:FigS2)B; lnRR = 0.111, CI = [0.083 – 0.138]) while 9.38 % towards females (lnRR = -0.094, CI = [-0.129 – (-0.068)]. The overall pattern for lnRR was more varied among the functional trait groups than for lnCVR, ranging from 1.2% (Hematology, lnRR = -0.013, CI = [-0.017 – (-0.007)] to 14.72 % (Immunology, lnRR = -0.159, CI = [-0.229 – (-0.089)] in female-based traits (Figure S\@ref(fig:FigS2)B) whereas in males-based traits from 1.8 % (Hearing, lnRR = 0.018, CI = [0.006 – 0.031]) to 25 % (Metabolism, lnRR = 0.223, CI = [0.110 – 0.335])  (Figure S\@ref(fig:FigS2)B).

```{r FigS2, fig.height = 10, fig.width = 8, fig.cap=" A) Differences in numbers of affected traits, in variance (lnCVR and lnVR) and means (lnRR), where there’s a significant difference between the sexes (i.e CI not overlapping zero), and where the sex bias is greater than 10% difference (regardless of significance). Panel B depicts results for the sex bias in those traits that differ between the sexes (second-order meta-analysis). Triangles represent sex bias in means (response ratio) and black circles differences in the coefficient of variation ratio (mean-adjusted variability). The orange-red bars represent trait groups with a female bias, blue-green bars male-biased traits."}
library(ggpubr)
FigS2b <- ggarrange(Metameta_FigS2_female.sig, Metameta_FigS2_male.sig,
  ncol = 2, nrow = 1, widths = c(1, 1.30), heights = c(1.2, 1.2))

@@

# end combination Figure 5
FigS2 <- ggarrange(malebias_FigS2_sigtraits, malebias_Fig2_over10, FigS2b, FigS2d, ncol = 1, nrow = 4, heights = c(2.5, 2.2, 2.1, 2), labels = c("A", " ", "B", " "))
FigS2
```

# Supplemental Tables
## Table 6.1

@@

## Table 6.2

@@

## Table 6.3

@@

## Table 6.4

@@

## Table 6.5

@@

## Table 6.6

>>>>

---
title: 'Sexual dimorphism in trait variability and its eco-evolutionary
  and statistical implications'
author: Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky, Robert  Brooks,
  Will Cornwell, Daniel Falster, Malgorzata Lagisz, Jeremy Mason, Alistair Senior,
  Daniel Noble, & Shinichi Nakagawa
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 6
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
  html_notebook:
    toc: yes
  bookdown::pdf_document2:
    toc: yes
    toc_depth: 6
    number_sections: true
    theme: flatly
  bookdown::word_document2:
    toc: yes
    toc_depth: 6
    number_sections: true
    theme: flatly
subtitle: Electronic Supporting File (doi:10.5281/zenodo.4146948)
always_allow_html: yes
---

@@

```{r FigG1, echo = FALSE, fig.width = 4, fig.height = 8, fig.cap="Workflow of data processing and meta-analysis [Figure 3 in manuscipt]"}

@@

```{r FigCor, fig.height=3, message = FALSE, fig.cap= "Mean-variance relationships (log(Mean) vs log(SD, standard deviation)) across all traits for males (A) and females (B).[Figure 1-figure supplement 1 in manuscript]"}

@@

# Meta-analysis results
## Preparing: Figure 4 
### Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. Preparation of subplot 3A

@@

## Main Manuscript: Figure 4

```{r Fig3, fig.cap = "Panel A shows the numbers of traits across functional groups that are either male-biased (blue-green) or female-biased (orange-red), as calculated in Figure 1.1D. Panel B shows effect sizes and 95% CI from separate meta-analysis for each functional group (step H in Figure1.1). Both panels represent results evaluated across all traits. Traits that are male biased are shown in blue, whereas female bias data is represented in orange.[Figure 4 in manuscript]"}
Fig4 <- ggarrange(malebias_Fig2_alltraits, Metameta_Fig3_alltraits,  nrow = 2, align = "v", heights = c(10, 9), labels = c("A", "B"))
Fig4
ggsave("Figure4.pdf", plot = Fig4, width = 6, height = 7)
```

@@

plot3.male.meta.CVR.b <- rbind(plot3.male.meta.CVR.b, add.row.hearing) #bind_rows

@@

overall4.male.sig <- rbind(lnCVR.ci, lnRR.ci) %>%  #bind_rows

@@

plot3.female.meta.CVR <- bind_rows(plot3.female.meta.CVR, plot3.female.meta.CVR.all) #bind_rows

@@

plot3.female.meta.CVR.b <- rbind(plot3.female.meta.CVR.b, add.row.hearing) #bind_rows

@@

overall4.female.sig <- rbind(lnCVR.ci, lnRR.ci) %>%  #bind_rows

@@

overall4.male.sigS <- rbind(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #bind_rows

@@

overall4.female.sigS <- rbind(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) # Note: bind_rows may need to be replaced by rbind

@@

overall4.male.sigS <- rbind(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) %>% 
  mutate(value = as.numeric(value), ci.low = as.numeric(ci.low), ci.high = as.numeric(ci.high))
# may require: 	
overall4.male.sigS$label <- "CI not overlapping zero"

@@

## Figure 5.1
This Figure is similar to Figure 4 in the main article except that lnVR has been added to the plot for comparison with lnCVR and lnRR.

```{r FigS1, fig.height = 8, fig.width = 7, fig.cap=" Numbers of either male (blue-green bars) or female (orange-red bars) biased traits  (Panel A) across functional groups, this time for lnCVR (left hand side), lnVR (middle) and lnRR (right hand side). Panel B shows effect sizes from separate meta-analysis for each functional group, and Panel C contains results of heterogeneity analyses. All three panels represent results evaluated across all traits.[Figure 4 – figure supplement 1 in manuscript]"}

@@

FigS1 <- ggarrange(malebias_FigS1_alltraits, Metameta_FigS1_alltraits, HeteroS1, nrow = 3, align = "v", heights = c(1.1, 1, 1), labels = c("A", "B", "C"))
FigS1
ggsave("Figure4_1.pdf", plot = FigS1, width = 6, height = 7)
```

@@

#### Methods
Among the 147 traits (i.e. with ‘first-order’ meta-analytic results), we found 31 traits to be significantly male-biased and 30 female-biased for lnCVR (47 male-biased and 39 female- biased for lnRR; 74 male and 55 female-biased for lnVR). We used this statistical significance to categorize traits into two groups: male-based and female-biased traits. For each set of these sex-biased traits, we conducted second-order meta-analyses with trait as the unit of analysis in the same way as those using lnRR, lnVR and lnCVR (Figure \@ref(fig:FigG1)J).  As sensitivity analyses, we also employed another way of separating male- and female-biased traits. We quantified traits that were more than 10% biased towards either of the sexes, regardless of significance (N[male-biased] = 31, & N[female-biased] = 36 for lnCVR,  N[male-biased] = 56, & N[female-biased] = 57 for lnVR , N[male-biased] = 42, & N[female-biased] = 33  for lnRR; see Figure 4 – figure supplement 1, Supplemental Table 3).

#### Results
Having established that different sex-biases were present across trait functional groups, we explored specific patterns in male and female bias by categorizing traits into either male or female-biased, and then, quantifying the degree of sex bias. To do this, we focused on only those traits that were significantly biased towards one sex (as indicated by “CI not overlapping 0” in Figure Figure 4 – figure supplement 2). Selecting significantly sex-biased traits reduced the number of traits substantially (Figure 4 – figure supplement 2A; compared to all traits, Figure 4 and Figure 4 – figure supplement 1), and returned similar numbers of traits with either significant male- or female-bias for lnCVR (Figure 4 – figure supplement 2A).

Looking only at traits with significant male-bias, trait variability tended to be higher in males than in females overall (11.77% male bias, lnCVR = 0.111, CI = [0.093 to 0.129]; compared to 8.3% female-bias, lnCVR = -0.087, CI = [-0.109 to (-0.064)]; Figure 4 – figure supplement 2B). The degree of sex-bias varied across the functional trait groups, ranging from 4.6 % (Physiology, lnCVR = -0.047, CI = [-0.064 to (-0.030)] to 15.13 % (Eye morphology, lnCVR = -0.164, CI = [-0.249 to (-0.250)] in females, and from 8.6 % (Hematology, lnCVR = 0.082, CI = [0.048 to 0.117]) and 16.66 % (Heart, lnCVR = 0.154, CI = [0.085 to 0.223]) in males (Figure 4 – figure supplement 2B).

To put sex biases (sexual dimorphisms) in variability into perspective, we also quantify the degree of sexual dimorphism in significantly sex-biased traits (i.e., the extent of sexual dimorphisms). There were more traits that were significantly biased in mean trait value in males compared to females (Figure 4 – figure supplement 2A, lnRR). These traits were 11.73 % biased towards males Figure 4 – figure supplement 2B; lnRR = 0.111, CI = [0.083 – 0.138]) while 9.38 % towards females (lnRR = -0.094, CI = [-0.129 – (-0.068)]. The overall pattern for lnRR was more varied among the functional trait groups than for lnCVR, ranging from 1.2% (Hematology, lnRR = -0.013, CI = [-0.017 – (-0.007)] to 14.72 % (Immunology, lnRR = -0.159, CI = [-0.229 – (-0.089)] in female-based traits Figure 4 – figure supplement 2B, whereas in males-based traits from 1.8 % (Hearing, lnRR = 0.018, CI = [0.006 – 0.031]) to 25 % (Metabolism, lnRR = 0.223, CI = [0.110 – 0.335])  (Figure 4 – figure supplement 2B).

```{r FigS2, fig.height = 10, fig.width = 8, fig.cap=" A) Differences in numbers of affected traits, in variance (lnCVR and lnVR) and means (lnRR), where there is a significant difference between the sexes (i.e CI not overlapping zero), and where the sex bias is greater than 10% difference (regardless of significance). Panel B depicts results for the sex bias in those traits that differ between the sexes (second-order meta-analysis). Triangles represent sex bias in means (response ratio) and black circles differences in the coefficient of variation ratio (mean-adjusted variability). The orange-red bars represent trait groups with a female bias, blue-green bars male-biased traits.[Figure 4 – figure supplement 2 in manuscript]"}
library(ggpubr)
FigS2b <- ggarrange(Metameta_FigS2_female.sig, Metameta_FigS2_male.sig,
  ncol = 2, nrow = 1, widths = c(1, 1.30), heights = c(1.2, 1.2))

@@

# end combination Figure 5
FigS2 <- ggarrange(malebias_FigS2_sigtraits, malebias_Fig2_over10, FigS2b, FigS2d, ncol = 1, nrow = 4, heights = c(2.5, 2.2, 2.1, 2), labels = c("A", " ", "B", " "))
FigS2
ggsave("Figure4_2.pdf", plot = FigS2, width = 6, height = 7)
```

# .Supplemental Tables
## Table 1

@@

## Table 2

@@

## Table 3

@@

## Table 4

@@

## Table 5

@@

## Table 6
